---
title: "Event Streaming"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Event Streaming}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(puppeteeR)
```

puppeteeR provides two ways to observe a running graph:

| Mechanism | Granularity | How to use |
|---|---|---|
| `on_step` | One callback per node, after it finishes | `config = list(on_step = fn)` |
| `on_event` | Two callbacks per node - before and after | `config = list(on_event = fn)` |
| `stream()` | One yielded value per node, after it finishes | `runner$stream()` generator |

`on_step` and `stream()` exist for simple progress tracking. `on_event` is the richer
replacement: it fires twice per node and carries structured data, making it suitable for
logging, timing, dashboards, and debugging.

## The event object

Every `on_event` call receives a single list with four fields:

| Field | Type | Description |
|---|---|---|
| `type` | character | `"node_start"` or `"node_end"` |
| `node` | character | Name of the node being executed |
| `iteration` | integer | Loop counter (1-based) |
| `data` | list or NULL | `NULL` on `node_start`; `list(updates = <named list>)` on `node_end` |

`data$updates` on `node_end` is exactly the named list that the node function returned -
the raw state updates before reducers have been applied.

## Basic usage

```{r basic}
schema <- workflow_state(
  result = list(default = "")
)

runner <- state_graph(schema) |>
  add_node("prepare", function(state, config) list(result = "prepared")) |>
  add_node("finalise", function(state, config) {
    list(result = paste(state$get("result"), "and finalised"))
  }) |>
  add_edge(START, "prepare") |>
  add_edge("prepare", "finalise") |>
  add_edge("finalise", END) |>
  compile()

runner$invoke(config = list(
  on_event = function(event) {
    cat(sprintf("[%s] %s (iter %d)\n", event$type, event$node, event$iteration))
  }
))
```

## Timing nodes

Because `node_start` fires before execution and `node_end` fires after, you can measure
wall-clock time for each node without modifying any node function:

```{r timing}
timings <- list()
start_times <- list()

runner$invoke(config = list(
  on_event = function(event) {
    if (event$type == "node_start") {
      start_times[[event$node]] <<- proc.time()[["elapsed"]]
    } else {
      elapsed <- proc.time()[["elapsed"]] - start_times[[event$node]]
      timings[[event$node]] <<- elapsed
    }
  }
))

str(timings)
```

## Inspecting state updates

`data$updates` on a `node_end` event lets you log exactly what each node changed without
adding `print()` calls inside node functions:

```{r updates}
schema2 <- workflow_state(
  count  = list(default = 0L),
  status = list(default = "idle")
)

runner2 <- state_graph(schema2) |>
  add_node("step_a", function(state, config) list(count = 1L, status = "running")) |>
  add_node("step_b", function(state, config) list(count = 2L, status = "done")) |>
  add_edge(START, "step_a") |>
  add_edge("step_a", "step_b") |>
  add_edge("step_b", END) |>
  compile()

runner2$invoke(config = list(
  on_event = function(event) {
    if (event$type == "node_end") {
      cat(sprintf("  %s returned: %s\n",
        event$node,
        paste(names(event$data$updates), collapse = ", ")
      ))
    }
  }
))
```

## Building an execution log

Collect events into a data frame for post-run analysis:

```{r log}
log_entries <- list()

runner2$invoke(config = list(
  on_event = function(event) {
    log_entries[[length(log_entries) + 1L]] <<- data.frame(
      type      = event$type,
      node      = event$node,
      iteration = event$iteration,
      stringsAsFactors = FALSE
    )
  }
))

do.call(rbind, log_entries)
```

## Combining on_event with on_step

`on_step` is kept for backwards compatibility and works alongside `on_event`. Both can be
supplied in the same config - `on_event` fires first (before and after the node),
then `on_step` fires once after the node completes:

```{r both}
runner$invoke(config = list(
  on_event = function(event) cat("event:", event$type, event$node, "\n"),
  on_step  = function(node, state) cat("step: ", node, "\n")
))
```

## Current limitation - tool-level events

`on_event` currently emits only node-level events. Two additional event types are planned
but not yet available:

| Planned type | Meaning |
|---|---|
| `tool_call` | An LLM inside the node called a tool |
| `tool_result` | The tool returned a result to the LLM |

These sub-node events require ellmer to expose an intercept hook on `ellmer::Chat` so
puppeteeR can observe individual tool invocations without wrapping the Chat object. That
hook does not exist yet. Once it is added upstream, `tool_call` and `tool_result` events
will be emitted automatically without any change to the `on_event` API.

Until then, if you need visibility into tool calls, the simplest workaround is to wrap
the tool function itself:

```{r workaround, eval = FALSE}
observed_search <- tool(
  fn = function(query) {
    cat("[tool_call] search:", query, "\n")
    result <- real_search(query)
    cat("[tool_result] search:", nchar(result), "chars\n")
    result
  },
  description = "Search the web",
  arguments = list(query = type_string("Search query"))
)
```

## See also

- `vignette("checkpointing")` - persisting state between runs
- `vignette("retry-policy")` - retrying nodes that fail transiently
- `?GraphRunner` - full reference for `$invoke()` and `$stream()`
