---
title: "Getting Started with puppeteeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with puppeteeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(puppeteeR)
```

puppeteeR is an LLM multi-agent orchestrator for R built on [ellmer](https://ellmer.tidyverse.org).
It lets you define directed graphs where each node is an R function (or an LLM agent) that reads
shared state, does work, and writes updates back. Think LangGraph, but idiomatic R.

## Core concepts

| Concept | What it is |
|---|---|
| `WorkflowState` | Mutable key-value store shared across all nodes |
| `StateGraph` | Builder for the graph (nodes + edges) |
| `GraphRunner` | Compiled, executable graph returned by `compile()` |
| `Agent` | Thin wrapper around an `ellmer::Chat` with identity + cost tracking |

## 1. Define state

Every graph has a **state schema** - a named set of channels, each with a default value and an
optional reducer that controls how updates merge.

```{r state}
schema <- workflow_state(
  input   = list(default = ""),
  result  = list(default = NULL),
  history = list(default = list(), reducer = reducer_append())
)
```

Three built-in reducers:

- `reducer_overwrite()` - default, replaces old value with new value
- `reducer_append()` - wraps new value in a list and appends to old list (great for message history)
- `reducer_merge()` - shallow-merges named lists with `modifyList()`

## 2. Build the graph

Nodes are plain R functions with the signature `function(state, config)`. They return a named list
of channel updates.

```{r graph}
runner <- state_graph(schema) |>
  add_node("process", function(state, config) {
    val <- toupper(state$get("input"))
    list(result = val, history = val)
  }) |>
  add_node("enrich", function(state, config) {
    prev <- state$get("result")
    list(result = paste0("[", prev, "]"), history = "enriched")
  }) |>
  add_edge(START, "process") |>
  add_edge("process", "enrich") |>
  add_edge("enrich", END) |>
  compile()
```

The pipe chain ends with `compile()`, which validates the graph and returns a `GraphRunner`.

## 3. Run it

```{r invoke}
final <- runner$invoke(list(input = "hello world"))

final$get("result")
final$get("history")
```

`invoke()` accepts an `initial_state` list that overrides channel defaults for this run.

## 4. Conditional routing

Return a key from a routing function to choose the next node dynamically.

```{r conditional}
schema2 <- workflow_state(
  n   = list(default = 0L),
  out = list(default = "")
)

runner2 <- state_graph(schema2) |>
  add_node("check", function(s, cfg) list()) |>
  add_node("big",   function(s, cfg) list(out = "big number")) |>
  add_node("small", function(s, cfg) list(out = "small number")) |>
  add_edge(START, "check") |>
  add_conditional_edge(
    from       = "check",
    routing_fn = function(s) if (s$get("n") > 10L) "big" else "small",
    route_map  = list(big = "big", small = "small")
  ) |>
  add_edge("big",   END) |>
  add_edge("small", END) |>
  compile()

runner2$invoke(list(n = 3L))$get("out")
runner2$invoke(list(n = 99L))$get("out")
```

## 5. Adding an LLM agent

Agents wrap `ellmer::Chat` objects. Pass them to `compile()` and access them inside nodes via
`config$agents`.

```{r agent, eval = FALSE}
library(ellmer)

researcher <- agent(
  name         = "researcher",
  chat         = chat_anthropic(model = "claude-haiku-4-5"),
  role         = "Senior researcher",
  instructions = "Give concise, factual answers."
)

schema3 <- workflow_state(
  query  = list(default = ""),
  answer = list(default = "")
)

runner3 <- state_graph(schema3) |>
  add_node("respond", function(state, config) {
    ans <- config$agents$researcher$chat(state$get("query"))
    list(answer = ans)
  }) |>
  add_edge(START, "respond") |>
  add_edge("respond", END) |>
  compile(agents = list(researcher = researcher))

result <- runner3$invoke(list(query = "What is the speed of light?"))
result$get("answer")
```

## 6. Automatic retries

LLM API calls occasionally fail with transient errors — dropped connections, rate limits,
temporary service issues. Every `agent()` retries failed calls automatically before raising
an error, so your pipeline keeps running without any extra code.

The defaults are 3 retries with a 5-second wait between attempts. Override per agent:

```{r retries, eval = FALSE}
# More patient agent for a slow or rate-limited endpoint
researcher <- agent(
  name         = "researcher",
  chat         = chat_anthropic(),
  instructions = "Be thorough.",
  max_retries  = 5L,   # up to 5 retries
  retry_wait   = 10    # 10 seconds between each
)

# Disable retries entirely (fail fast)
fast_agent <- agent("fast", chat_anthropic(), max_retries = 0L)
```

Each retry attempt logs a warning with the error message and countdown, so you can see what
is happening without the pipeline silently hanging.

For retrying entire **graph nodes** on error — useful when a node calls an external API that
may fail — see `vignette("retry-policy")` for the `retry_policy()` interface.

## 7. Streaming execution

`stream()` returns a `coro` generator that yields after each node - useful for showing progress.

```{r stream, eval = FALSE}
gen <- runner3$stream(list(query = "Explain quantum entanglement briefly."))
coro::loop(for (step in gen) {
  cat("Node:", step$node, "| iteration:", step$iteration, "\n")
})
```

## Next steps

- **Custom graphs**: conditional loops, multi-agent collaboration → `vignette("custom-graphs")`
- **Convenience workflows**: one-liner sequential / supervisor / debate → `vignette("workflows")`
- **Checkpointing**: resume interrupted runs, human-in-the-loop → `vignette("checkpointing")`
- **Retry policies**: per-node automatic retries with backoff → `vignette("retry-policy")`
- **Visualization**: render graph diagrams → `vignette("visualization")`
