---
title: "Node Retry Policies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Node Retry Policies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(puppeteeR)
```

puppeteeR has two independent retry mechanisms:

| Mechanism | What it retries | Where it is configured |
|---|---|---|
| **Agent-level** (`max_retries`, `retry_wait`) | A single LLM API call inside `agent$chat()` | `agent()` / `Agent$new()` |
| **Node-level** (`retry_policy()`) | An entire node function execution | `add_node(..., retry = ...)` |

Agent-level retries handle transient API errors automatically for every LLM call. Node-level
retry policies are for when the node itself is unreliable — an external HTTP call, a database
query, or any fallible operation that a node performs beyond just calling an LLM.

## Creating a retry policy

`retry_policy()` takes three arguments:

```{r policy}
# Retry up to 3 times (1 initial attempt + 2 retries), wait 1 second between attempts
rp <- retry_policy(max_attempts = 3L, wait_seconds = 1, backoff = 1)
rp
```

| Argument | Type | Default | Meaning |
|---|---|---|---|
| `max_attempts` | integer ≥ 2 | `3L` | Total number of attempts including the first |
| `wait_seconds` | number ≥ 0 | `1` | Seconds to wait before the next attempt |
| `backoff` | number > 0 | `1` | Multiplier applied to `wait_seconds` after each failure |

With `backoff = 1` the wait is constant. With `backoff = 2` the wait doubles each time:
1 s → 2 s → 4 s → …

## Basic usage

Pass a `retry_policy` to `add_node()` via the `retry` argument:

```{r basic}
# Simulate a node that fails twice then succeeds
attempt_n <- 0L

flaky_node <- function(state, config) {
  attempt_n <<- attempt_n + 1L
  if (attempt_n < 3L) stop("transient error")
  list(result = paste0("succeeded on attempt ", attempt_n))
}

schema <- workflow_state(result = list(default = ""))

runner <- state_graph(schema) |>
  add_node(
    "fetch",
    flaky_node,
    retry = retry_policy(max_attempts = 3L, wait_seconds = 0)
  ) |>
  add_edge(START, "fetch") |>
  add_edge("fetch", END) |>
  compile()

final <- runner$invoke()
final$get("result")
attempt_n
```

The node failed twice and succeeded on the third attempt. From the caller's perspective,
`invoke()` returned normally — the retries were invisible.

## Exhausting all attempts

When all attempts fail, the runner re-throws the last error with a message that includes the
node name and attempt count:

```{r exhausted, error = TRUE}
always_fail <- function(state, config) stop("permanent failure")

schema2 <- workflow_state(x = list(default = 0L))

runner2 <- state_graph(schema2) |>
  add_node(
    "broken",
    always_fail,
    retry = retry_policy(max_attempts = 3L, wait_seconds = 0)
  ) |>
  add_edge(START, "broken") |>
  add_edge("broken", END) |>
  compile()

tryCatch(
  runner2$invoke(),
  error = function(e) message(conditionMessage(e))
)
```

## Exponential backoff

For external API calls, exponential backoff reduces load on a struggling service. Set
`backoff = 2` to double the wait after each failure:

```{r backoff, eval = FALSE}
runner <- state_graph(workflow_state(data = list(default = ""))) |>
  add_node(
    "call_api",
    function(state, config) {
      response <- httr2::request("https://api.example.com/data") |>
        httr2::req_perform() |>
        httr2::resp_body_string()
      list(data = response)
    },
    retry = retry_policy(max_attempts = 4L, wait_seconds = 1, backoff = 2)
    # Waits: 1s → 2s → 4s between attempts
  ) |>
  add_edge(START, "call_api") |>
  add_edge("call_api", END) |>
  compile()
```

## Node retry vs. agent-level retry

The two mechanisms target different failure modes and compose independently:

```{r comparison, eval = FALSE}
library(ellmer)

# Agent-level: retries the LLM API call if the HTTP request fails
researcher <- agent(
  "researcher",
  chat_anthropic(),
  max_retries = 3L,    # 4 total LLM call attempts
  retry_wait  = 5      # 5 seconds between LLM call retries
)

# Node-level: retries the entire node function if it throws any error
runner <- state_graph(workflow_state(result = list(default = ""))) |>
  add_node(
    "research",
    function(state, config) {
      # This node does two things: fetch data from an external API, then call the LLM.
      # If either fails, the node-level policy retries the whole thing.
      raw <- fetch_external_data(state$get("query"))   # may fail with network errors
      result <- config$agents$researcher$chat(raw)     # may fail with API errors (caught by agent retry first)
      list(result = result)
    },
    retry = retry_policy(max_attempts = 2L, wait_seconds = 30)
  ) |>
  add_edge(START, "research") |>
  add_edge("research", END) |>
  compile(agents = list(researcher = researcher))
```

When the node runs:

1. `fetch_external_data()` fails → node-level retry fires after 30 s.
2. `researcher$chat()` fails → agent-level retry fires (up to 4 attempts, 5 s apart). If all
   agent retries fail, the error propagates out of `$chat()` and the node-level policy retries
   the whole node.

The two policies stack: agent-level retry guards individual LLM calls; node-level retry guards
the entire operation including any surrounding code.

## Disabling retries for a node

Omit the `retry` argument (or pass `NULL`) to run a node without any retry logic. In that
case a single failure immediately propagates as an error from `invoke()`:

```{r no-retry, eval = FALSE}
state_graph(schema) |>
  add_node("fast_fail", function(state, config) stop("oops"))  # retry = NULL (default)
```

## Combining retry policy with checkpointing

Node-level retries happen within a single `invoke()` call — the checkpointer does not see
individual retry attempts, only the final outcome of each node. This means:

- A successful retry → checkpoint is saved as if the node succeeded first time.
- All retries exhausted → the error propagates; no checkpoint is written for that node.

On the next `invoke()` with the same `thread_id`, the runner resumes from the last successful
checkpoint and retries the failing node from scratch.

```{r checkpoint-combo, eval = FALSE}
cp <- rds_checkpointer(path = "checkpoints/")

schema <- workflow_state(
  data   = list(default = ""),
  result = list(default = "")
)

runner <- state_graph(schema) |>
  add_node("fetch", function(state, config) {
    if (nzchar(state$get("data"))) return(list())   # idempotent: skip if already fetched
    list(data = fetch_from_api())                   # may fail; checkpoint saved after success
  }) |>
  add_node(
    "process",
    function(state, config) {
      list(result = config$agents$llm$chat(state$get("data")))
    },
    retry = retry_policy(max_attempts = 3L, wait_seconds = 5, backoff = 2)
  ) |>
  add_edge(START, "fetch") |>
  add_edge("fetch", "process") |>
  add_edge("process", END) |>
  compile(agents = list(llm = my_agent), checkpointer = cp)

tryCatch(
  runner$invoke(config = list(thread_id = "job-1")),
  error = function(e) message("Failed: ", conditionMessage(e))
)

# On re-run, fetch is skipped (idempotent guard) and process is retried fresh
runner$invoke(config = list(thread_id = "job-1"))
```

## Validation

`retry_policy()` validates its arguments eagerly:

```{r validation, error = TRUE}
retry_policy(max_attempts = 1L)    # must be >= 2
```

```{r validation2, error = TRUE}
retry_policy(wait_seconds = -1)    # must be non-negative
```

```{r validation3, error = TRUE}
retry_policy(backoff = 0)          # must be positive
```

Passing a non-`retry_policy` object to `add_node()` is also rejected:

```{r validation4, error = TRUE}
schema3 <- workflow_state(x = list(default = 0L))
state_graph(schema3) |>
  add_node("n", function(state, config) list(), retry = list(max_attempts = 3L))
```

## See also

- `?retry_policy` — reference documentation
- `vignette("checkpointing")` — resuming interrupted runs with checkpointers
- `vignette("best-practices")` — agent-level retry tuning and context management
