---
title: "State Channels and Reducers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{State Channels and Reducers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(puppeteeR)
```

`WorkflowState` is the shared memory that all nodes in a graph read from and write to. Think
of it as a whiteboard: nodes read whatever is on it, write their results back, and the next
node picks up from there.

This vignette explains how that whiteboard is structured, what reducers are and why they
matter, and how to design a schema that fits your workflow.

## Two kinds of memory

puppeteeR has two memory systems that coexist independently:

**WorkflowState channels** are explicit, structured, and inspectable. Every node reads from
them with `state$get("name")` and writes to them by returning a named list. You define them
up front in a schema.

**ellmer Chat history** is implicit and per-agent. Every `agent$chat("...")` call appends
the exchange to that agent's internal conversation. The agent accumulates context at the LLM
level automatically, regardless of what WorkflowState contains.

These two systems do not synchronise. Resetting or restoring a WorkflowState from a
checkpoint does not touch any agent's Chat history. For most workflows this is fine; for
workflows that fork or replay, you may need to reset agent Chat objects manually:

```r
agent$chat_object$set_turns(list())
```

## Channels

A channel is a named slot in the state. You declare all channels up front in a schema:

```{r schema}
ws <- workflow_state(
  messages = list(default = list(), reducer = reducer_append()),
  status   = list(default = "pending"),
  metadata = list(default = list(), reducer = reducer_merge())
)
```

Each channel needs two things:

- `default` — the value it holds before any node has written to it.
- `reducer` — a `function(old, new)` that controls what happens when a node writes a new
  value. If you omit `reducer`, puppeteeR picks a sensible default automatically (more on
  this below).

Channel names are completely free — `status`, `my_counter`, `judge_verdict` — any valid R
name works. The engine has no reserved names. The only rule: the name you declare in the
schema must be the exact same name used in `state$get()` and in node return values.

## Reducers

A reducer is a two-argument function `function(old, new)` called every time a node writes
to a channel. It decides how the incoming value is merged with the current one.

The simplest possible reducer just replaces the old value:

```r
function(old, new) new
```

A more interesting one appends:

```r
function(old, new) c(old, list(new))
```

puppeteeR ships four built-in reducers that cover almost every use case.

## Built-in reducers

### `reducer_overwrite()` — replace

```{r overwrite}
r <- reducer_overwrite()
r("old value", "new value")
```

Discards `old` and returns `new`. This is the **default for non-list channels** (i.e.
anything whose `default` is not a `list()`). Use it for:

- Status flags and routing signals (`"pending"`, `"done"`, `"approved"`)
- The current version of a document
- Counters and numeric accumulators where you compute the new value yourself

### `reducer_append()` — accumulate

```{r append}
r <- reducer_append()
r(list("first"), "second")
```

Wraps `new` in `list()` and concatenates it to `old`. Use it for any channel that should
grow over time: conversation history, collected results, log entries.

> **Warning.** `reducer_append()` grows without bound. In a long-running workflow, an
> append channel fed to every LLM call will eventually exceed the context window and cause
> connection errors. Use `reducer_last_n(n)` instead when passing the channel to agents
> (see below).

### `reducer_last_n(n)` — sliding window

```{r last_n}
r <- reducer_last_n(3L)
r(list("a", "b", "c"), "d")   # drops "a"
```

Appends then trims to the most recent `n` entries. This is also the **automatic default for
list channels** (those with `default = list()`) when no `reducer` is specified — with a
window of 20. See the pitfall section below.

### `reducer_merge()` — shallow merge

```{r merge}
r <- reducer_merge()
r(list(model = "haiku", temp = 0.7), list(temp = 0.2, seed = 42L))
```

Uses `modifyList()`: keys in `new` overwrite matching keys in `old`; keys absent from `new`
are preserved. Use it for configuration or metadata objects that are updated piecemeal.

## The default reducer pitfall

When you declare a list channel without a `reducer`, puppeteeR assigns `reducer_last_n(20)`
automatically:

```r
# These two are equivalent:
plan = list(default = list())
plan = list(default = list(), reducer = reducer_last_n(20))
```

This is almost never what you want for a structured data channel. Consider a `plan` channel
that holds a list of step objects:

```r
planner_fn <- function(state, config) {
  steps <- parse_plan(response)          # list of 6 step objects
  list(plan = steps)                     # writes the list of 6 steps
}
```

With `reducer_last_n(20)`, this write does not *replace* `plan` with the 6 steps — it
*appends* the entire `steps` list as a single element. The result is
`list(list(step1, step2, ...))`. When the dispatcher later reads `plan[[1]]`, it gets the
whole plan, not step 1.

**Rule:** any list channel that is meant to be replaced wholesale must declare
`reducer = reducer_overwrite()` explicitly:

```r
plan = list(default = list(), reducer = reducer_overwrite())
```

Only declare `reducer_append()` or `reducer_last_n()` for channels that genuinely
accumulate entries over time.

## Designing your schema

### Append for history, overwrite for signals

The most common pattern:

```{r pattern1, eval = FALSE}
state_schema <- workflow_state(
  messages      = list(default = list(), reducer = reducer_append()),
  current_route = list(default = "")
)
```

`messages` accumulates the full conversation. `current_route` is a transient routing
signal — only the latest value matters, so overwrite is correct.

### Separate the current draft from history

When agents revise output iteratively, separate the live draft from the audit trail:

```{r pattern2, eval = FALSE}
state_schema <- workflow_state(
  messages     = list(default = list(), reducer = reducer_append()),
  latest_draft = list(default = ""),
  revision_n   = list(default = 0L)
)
```

Worker nodes write their output to `latest_draft` (overwrite — always the current version).
The advisor reads `latest_draft`, not `messages`. The full history in `messages` serves as
an audit trail without interfering with evaluation.

### Dedicated channel per routing signal

Never route based on the last item in an accumulating channel. Use a dedicated overwrite
channel for every routing decision:

```{r pattern3, eval = FALSE}
state_schema <- workflow_state(
  messages      = list(default = list(), reducer = reducer_append()),
  judge_verdict = list(default = "continue")
)
```

The judge node writes `list(judge_verdict = "done")`. The conditional edge reads it
directly:

```r
g$add_conditional_edge("judge", function(state) state$get("judge_verdict"), route_map)
```

No text-scanning the last message, no fragile substring matching. If any future node
appends to `messages` after the judge runs, routing is unaffected.

## What the convenience workflows pass to agents

The `messages` channel accumulates the full conversation, but the built-in node functions
do not all pass the full history to their LLM calls:

| Workflow | What the node passes to the LLM |
|---|---|
| `sequential_workflow` workers | Last message only |
| `supervisor_workflow` manager | Full history concatenated |
| `supervisor_workflow` workers | Full history concatenated |
| `debate_workflow` debaters | Full history concatenated |
| `debate_workflow` judge | Full history concatenated |

When building custom graphs, pass context explicitly in your node function:

```{r full-context, eval = FALSE}
writer_fn <- function(state, config) {
  msgs    <- state$get("messages")
  context <- paste(vapply(msgs, as.character, character(1)), collapse = "\n")
  list(messages = config$agents$writer$chat(context))
}
```

## Custom reducers

Any `function(old, new)` works as a reducer. A common pattern is to wrap the inner
function in an outer factory function so it can take parameters:

```{r custom, eval = FALSE}
reducer_max <- function() {
  function(old, new) max(old, new, na.rm = TRUE)
}

state_schema <- workflow_state(
  messages   = list(default = list(), reducer = reducer_last_n(10L)),
  best_score = list(default = 0,      reducer = reducer_max())
)
```

The factory pattern (`reducer_max <- function() { ... }`) is not required for zero-argument
reducers, but it is consistent with the built-in style and makes the schema read cleanly:
`reducer = reducer_max()` vs `reducer = function(old, new) max(old, new, na.rm = TRUE)`.

## Snapshot and restore

`$snapshot()` returns a plain named list — a point-in-time copy of all channel values.
This is what checkpointers persist, and what `$restore()` reads back:

```{r snapshot, eval = FALSE}
ws <- workflow_state(
  messages = list(default = list(), reducer = reducer_append()),
  status   = list(default = "pending")
)

ws$update(list(messages = "hello", status = "running"))
snap <- ws$snapshot()   # plain list — safe to saveRDS(), serialize(), etc.

ws$update(list(messages = "world"))
ws$restore(snap)        # rolls back to the post-"hello" state
ws$get("messages")      # list("hello") — "world" is gone
```

`$restore()` bypasses reducers and directly overwrites channel values with the snapshot
contents. This is intentional: restoration must reproduce the exact prior state, not merge
into it.

## Schema reference for the built-in workflows

The schemas below are what the convenience constructors use internally. They illustrate
the design principles above in practice.

### `advisor_workflow()`

```{r advisor-schema, eval = FALSE}
state_schema <- workflow_state(
  messages         = list(default = list(), reducer = reducer_append()),
  latest_draft     = list(default = ""),         # current worker output — overwrite
  advisor_feedback = list(default = ""),         # advisor's revision notes — overwrite
  advisor_verdict  = list(default = "revise"),   # routing signal — overwrite
  revision_n       = list(default = 0L)          # revision counter — overwrite
)
```

Graph: `START → worker → advisor → approved → END`
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;`↑______________revise___|`

### `planner_workflow()`

```{r planner-schema, eval = FALSE}
state_schema <- workflow_state(
  messages            = list(default = list(), reducer = reducer_append()),
  plan                = list(default = list(), reducer = reducer_overwrite()),  # replaced wholesale each planner turn
  plan_index          = list(default = 0L),              # dispatcher cursor — overwrite
  current_instruction = list(default = ""),              # active step — overwrite
  current_worker      = list(default = ""),              # active step worker — overwrite
  results             = list(default = list(), reducer = reducer_append()),
  evaluator_verdict   = list(default = ""),              # routing signal — overwrite
  replan_count        = list(default = 0L)               # overwrite
)
```

`plan` uses `reducer_overwrite()` because the planner replaces the entire plan on each
turn — it is not accumulating steps, it is issuing a new set of them. Without the explicit
reducer, `reducer_last_n(20)` would wrap the new plan as a single nested element, breaking
the dispatcher.

Graph: `START → planner → dispatcher → workers (loop) → evaluator → done → END`
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;`↑_______________replan___|`
