---
title: "Visualizing Graphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing Graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(puppeteeR)
```

puppeteeR can render your graph structure before execution - useful for sanity-checking topology,
sharing diagrams, and documentation. Three engines are supported.

## Build a sample graph

```{r build-graph}
schema <- workflow_state(
  input    = list(default = ""),
  research = list(default = ""),
  draft    = list(default = ""),
  status   = list(default = "pending")
)

g <- state_graph(schema) |>
  add_node("research", function(s, cfg) list()) |>
  add_node("write",    function(s, cfg) list()) |>
  add_node("review",   function(s, cfg) list()) |>
  add_node("revise",   function(s, cfg) list()) |>
  add_edge(START, "research") |>
  add_edge("research", "write") |>
  add_edge("write", "review") |>
  add_conditional_edge(
    "review",
    routing_fn = function(s) if (s$get("status") == "approved") "done" else "needs_work",
    route_map  = list(done = END, needs_work = "revise")
  ) |>
  add_edge("revise", "write")
```

## Mermaid (text output, zero dependencies)

`as_mermaid()` returns a Mermaid flowchart string. Paste it into [mermaid.live](https://mermaid.live)
or embed it in any Markdown that supports Mermaid fences.

```{r mermaid}
cat(g$as_mermaid())
```

## DOT / Graphviz (requires `DiagrammeR`)

`as_dot()` returns a Graphviz DOT string. `visualize("dot")` renders it as an interactive widget.

```{r dot-source}
cat(g$as_dot())
```

```{r dot-render, eval = FALSE}
# Renders in RStudio Viewer / HTML output
g$visualize("dot")
```

Node colours:

- **Blue** (`#E8F0FE`) - regular nodes
- **Yellow** (`#FFF3CD`) - nodes with conditional outgoing edges
- **Dark** - START and END sentinels

## visNetwork (interactive, requires `visNetwork`)

Produces a pan-and-zoom interactive network widget. Solid lines = fixed edges,
dashed lines = conditional edges (labelled with the routing key).

```{r visnetwork, eval = FALSE}
g$visualize("visnetwork")
```

## Export to file

Export the DOT diagram to SVG or PNG (requires `DiagrammeR`, `DiagrammeRsvg`; PNG also needs `rsvg`).

```{r export, eval = FALSE}
g$export_diagram("workflow.svg")
g$export_diagram("workflow.png", width = 1200L, height = 800L)
```

## Visualizing compiled runners

Visualization works on `StateGraph` objects (before `compile()`). To inspect a compiled runner's
structure, keep a reference to the graph object:

```{r keep-graph, eval = FALSE}
g <- state_graph(schema) |>
  add_node("a", function(s, cfg) list()) |>
  add_edge(START, "a") |>
  add_edge("a", END)

runner <- g |> compile()

# g is still available for visualization
g$visualize("dot")
```

## Print summary

The `print()` method gives a quick text summary of node and edge counts:

```{r print}
g
```
