---
title: "Convenience Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Convenience Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(puppeteeR)
```

puppeteeR ships three one-liner constructors for common multi-agent patterns. Each returns a
compiled `GraphRunner` ready to invoke.

## `sequential_workflow()` — linear chain

Agents run in order. Each agent receives the previous agent's response as its input.

```{r sequential, eval = FALSE}
library(ellmer)

runner <- sequential_workflow(list(
  drafter  = agent("drafter",  chat_anthropic(),
                   instructions = "Write a first draft of the requested content."),
  reviewer = agent("reviewer", chat_anthropic(),
                   instructions = "Review the draft and suggest improvements."),
  polisher = agent("polisher", chat_anthropic(),
                   instructions = "Incorporate the review and produce a polished final version.")
))

result <- runner$invoke(list(messages = list("Write a short blog post about tidy data.")))

# All messages (input + all agent responses)
result$get("messages")
```

The default state schema has a single `messages` channel with `reducer_append()`. Each agent's
response is appended, so the final state contains the full conversation chain.

You can supply a custom schema if you need additional channels:

```{r sequential-custom, eval = FALSE}
my_schema <- workflow_state(
  messages = list(default = list(), reducer = reducer_append()),
  topic    = list(default = "")
)

runner2 <- sequential_workflow(
  agents       = list(
    researcher = agent("researcher", chat_anthropic()),
    writer     = agent("writer",     chat_anthropic())
  ),
  state_schema = my_schema
)

result2 <- runner2$invoke(list(
  topic    = "reinforcement learning",
  messages = list("Write about reinforcement learning.")
))
```

## `supervisor_workflow()` — hub-and-spoke

A manager agent directs work to a pool of worker agents. After each worker finishes, control
returns to the manager. The manager ends the workflow by replying `"DONE"`.

```{r supervisor, eval = FALSE}
runner <- supervisor_workflow(
  manager = agent(
    "manager", chat_anthropic(),
    instructions = paste0(
      "You coordinate a team. Available workers: 'researcher', 'writer'.\n",
      "Reply with exactly one worker name to delegate, or 'DONE' when finished."
    )
  ),
  workers = list(
    researcher = agent("researcher", chat_anthropic(),
                       instructions = "Research the given topic and return bullet-point notes."),
    writer     = agent("writer",     chat_anthropic(),
                       instructions = "Turn research notes into a readable article.")
  ),
  max_rounds = 6L
)

result <- runner$invoke(list(messages = list("Produce an article on quantum computing.")))
result$get("messages")
```

**How routing works**: after each manager turn, the supervisor searches the manager's response text
for a worker name. If none is found it routes to `"DONE"`. The `max_rounds` safety valve limits
total manager turns.

**Tip**: prompt the manager explicitly to name exactly one worker per turn. Ambiguous responses
fall through to `"DONE"`.

## `debate_workflow()` — round-robin with optional judge

Agents take turns responding to each other in round-robin order. Useful for exploring multiple
perspectives on a question, stress-testing an argument, or creative brainstorming.

### Without a judge

Runs for a fixed number of rounds then stops.

```{r debate-no-judge, eval = FALSE}
runner <- debate_workflow(
  agents = list(
    pro  = agent("pro",  chat_anthropic(),
                 instructions = "Argue in favour of the proposition."),
    con  = agent("con",  chat_anthropic(),
                 instructions = "Argue against the proposition.")
  ),
  max_rounds = 3L
)

result <- runner$invoke(list(
  messages = list("Proposition: R is better than Python for data science.")
))

# Print the debate transcript
for (msg in result$get("messages")) cat("---\n", as.character(msg), "\n")
```

### With a judge

After each full round, a judge agent decides whether the debate is settled. The judge should reply
`"continue"` or `"done"`.

```{r debate-judge, eval = FALSE}
runner <- debate_workflow(
  agents = list(
    optimist  = agent("optimist",  chat_anthropic(),
                      instructions = "Always highlight the positive side."),
    pessimist = agent("pessimist", chat_anthropic(),
                      instructions = "Always highlight risks and downsides.")
  ),
  judge = agent(
    "judge", chat_anthropic(),
    instructions = paste0(
      "Read the debate so far. If both sides have made their key points ",
      "and the debate is productive, reply 'done'. Otherwise reply 'continue'."
    )
  ),
  max_rounds = 5L
)

result <- runner$invoke(list(messages = list("Should we adopt AI in healthcare?")))
```

## Comparison

| Workflow | When to use |
|---|---|
| `sequential_workflow()` | Fixed pipeline: each step refines the previous output |
| `supervisor_workflow()` | Dynamic delegation: manager decides which specialist to call |
| `debate_workflow()` | Multiple perspectives: agents challenge each other |

For anything more complex — handoffs, parallel fans, custom state channels — build the graph
directly with `state_graph()`. See `vignette("custom-graphs")`.
