---
title: "Convenience Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Convenience Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(puppeteeR)
```

puppeteeR ships five one-liner constructors for common multi-agent patterns. Each returns a
compiled `GraphRunner` ready to invoke.

## `sequential_workflow()` - linear chain

Agents run in order. Each agent receives the previous agent's response as its input.

```{r sequential, eval = FALSE}
library(ellmer)

runner <- sequential_workflow(list(
  drafter  = agent("drafter",  chat_anthropic(),
                   instructions = "Write a first draft of the requested content."),
  reviewer = agent("reviewer", chat_anthropic(),
                   instructions = "Review the draft and suggest improvements."),
  polisher = agent("polisher", chat_anthropic(),
                   instructions = "Incorporate the review and produce a polished final version.")
))

result <- runner$invoke(list(messages = list("Write a short blog post about tidy data.")))

# All messages (input + all agent responses)
result$get("messages")
```

The default state schema has a single `messages` channel with `reducer_append()`. Each agent's
response is appended, so the final state contains the full conversation chain — one entry per
turn (initial prompt + one per agent).

```{r sequential-inspect, eval = FALSE}
msgs <- result$get("messages")
length(msgs)      # 4: initial prompt + drafter + reviewer + polisher

# Print the transcript
for (msg in msgs) cat("---\n", as.character(msg), "\n\n")
```

**Note on `state_schema`**: the `state_schema` parameter exists but is of limited use for
`sequential_workflow` because the built-in node functions only ever read and write the
`messages` channel. Any extra channels you add to the schema are invisible to the agents.
If you need agents that read from or write to custom channels, use `state_graph()` directly
and define your own node functions. See `vignette("custom-graphs")`.

## `supervisor_workflow()` - hub-and-spoke

A manager agent directs work to a pool of worker agents. After each worker finishes, control
returns to the manager. The manager ends the workflow by replying `"DONE"`.

```{r supervisor, eval = FALSE}
runner <- supervisor_workflow(
  manager = agent(
    "manager", chat_anthropic(),
    instructions = paste0(
      "You coordinate a team. Available workers: 'researcher', 'writer'.\n",
      "Reply with exactly one worker name to delegate, or 'DONE' when finished."
    )
  ),
  workers = list(
    researcher = agent("researcher", chat_anthropic(),
                       instructions = "Research the given topic and return bullet-point notes."),
    writer     = agent("writer",     chat_anthropic(),
                       instructions = "Turn research notes into a readable article.")
  ),
  max_rounds = 10L
)

result <- runner$invoke(list(messages = list("Produce an article on quantum computing.")))
result$get("messages")
```

**How routing works**: after each manager turn, the supervisor searches the manager's response text
for a worker name. If none is found it routes to `"DONE"`.

**`max_rounds` counts total node executions** — manager turns and worker turns combined. With 2
workers, a full delegation cycle (manager → worker → manager) consumes 3 turns. Set `max_rounds`
generously: `10L` gives room for roughly 3 full delegations plus a `"DONE"` turn.

**Tip**: prompt the manager explicitly to name exactly one worker per turn. Ambiguous responses
fall through to `"DONE"`.

## `debate_workflow()` - round-robin with optional judge

Agents take turns responding to each other in round-robin order. Useful for exploring multiple
perspectives on a question, stress-testing an argument, or creative brainstorming.

### Without a judge

Runs for a fixed number of rounds then stops.

```{r debate-no-judge, eval = FALSE}
runner <- debate_workflow(
  agents = list(
    pro  = agent("pro",  chat_anthropic(),
                 instructions = "Argue in favour of the proposition."),
    con  = agent("con",  chat_anthropic(),
                 instructions = "Argue against the proposition.")
  ),
  max_rounds = 3L
)

result <- runner$invoke(list(
  messages = list("Proposition: R is better than Python for data science.")
))

# Print the debate transcript
for (msg in result$get("messages")) cat("---\n", as.character(msg), "\n")
```

### With a judge

After each full round, a judge agent decides whether the debate is settled. The routing
checks whether the judge's response contains the word `"done"` (case-insensitive) anywhere
— if found, the workflow ends; otherwise it continues. The judge can write a full response
as long as it ends with `done` or `continue`.

```{r debate-judge, eval = FALSE}
runner <- debate_workflow(
  agents = list(
    optimist  = agent("optimist",  chat_anthropic(),
                      instructions = "Always highlight the positive side. Max 100 words."),
    pessimist = agent("pessimist", chat_anthropic(),
                      instructions = "Always highlight risks and downsides. Max 100 words.")
  ),
  judge = agent(
    "judge", chat_anthropic(),
    instructions = paste0(
      "You are moderating a debate. After each round, decide whether it has reached ",
      "a natural conclusion.\n\n",
      "If both sides have made their key points, write a brief summary and end your ",
      "response with the single word: done\n\n",
      "If the debate should continue, write a brief observation and end your ",
      "response with the single word: continue"
    )
  ),
  max_rounds   = 10L,
  state_schema = workflow_state(
    messages      = list(default = list(), reducer = reducer_last_n(8L)),
    judge_verdict = list(default = "continue")
  )
)

result <- runner$invoke(list(messages = list("Should we adopt AI in healthcare?")))

# Full transcript including judge summaries
for (msg in result$get("messages")) cat("---\n", as.character(msg), "\n\n")

# "done" means the judge ended it; "continue" means max_rounds fired first
result$get("judge_verdict")
```

**Custom `state_schema` and the `output` channel**: when you supply a `state_schema`, only
the channels you declare exist. The `output` channel is optional — the workflow runs fine
without it, but `result$output()` will error because there is no nominated output channel.
Add `output = list(default = "")` to your schema to re-enable it:

```{r debate-judge-output, eval = FALSE}
state_schema = workflow_state(
  messages      = list(default = list(), reducer = reducer_last_n(8L)),
  judge_verdict = list(default = "continue"),
  output        = list(default = "")   # enables result$output()
)
```

If you omit `output`, read the final judge response directly:

```{r debate-judge-no-output, eval = FALSE}
msgs <- result$get("messages")
as.character(msgs[[length(msgs)]])  # last message is the judge's closing summary
```

**`max_rounds` with a judge**: each round consumes 3 node executions (optimist + pessimist +
judge), but the internal termination guard is `max_turns(max_rounds * 2)` — it does not
account for judge turns. Set `max_rounds` higher than you need to ensure the judge gets
enough opportunities to run.

## `advisor_workflow()` — worker + advisor feedback loop

A cheap worker agent produces output; a higher-capability advisor agent evaluates it and
either approves it or sends it back with specific feedback. This pattern minimises cost by
running the expensive model only for evaluation, not for generation.

```{r advisor, eval = FALSE}
runner <- advisor_workflow(
  worker  = agent(
    "writer", chat_anthropic(model = "claude-haiku-4-5-20251001"),
    instructions = "Write clearly and concisely."
  ),
  advisor = agent(
    "advisor", chat_anthropic(model = "claude-opus-4-6"),
    instructions = paste0(
      "You are a strict editor. If the response is accurate, well-structured, and concise, ",
      "reply 'approved'. Otherwise reply 'revise: <specific feedback>'."
    )
  ),
  max_revisions = 3L
)

result <- runner$invoke(list(messages = list("Explain what a closure is in R.")))

result$get("latest_draft")   # final approved text
result$get("revision_n")     # number of revision cycles needed
result$get("messages")       # full audit trail: task, drafts, advisor verdicts
```

**How routing works**: after each advisor turn, the workflow reads the `advisor_verdict`
channel (`"approved"` or `"revise"`). If approved, execution ends. If revision is requested,
the worker is called again with the original task plus the advisor's feedback appended.

**State channels**: `latest_draft` (overwrite) always holds the current version — the advisor
evaluates this, not the full message history. `advisor_feedback` (overwrite) carries the most
recent revision notes. `messages` (append) is a full audit trail of every draft and verdict.

**Tip**: the routing checks whether the advisor's response starts with `"approved"` (case
insensitive). *Any* response that does not start with `"approved"` triggers a revision —
there is no special `"revise:"` keyword required for routing. However, the recommended
format `"revise: <feedback>"` improves feedback quality: the node strips the `"revise: "`
prefix and passes the remainder to the worker as targeted feedback. A response like `"needs
more examples"` also triggers revision but is passed to the worker verbatim as the feedback
string.

## `planner_workflow()` — Opus plans, Haiku executes

Separates expensive planning and evaluation (Opus-tier) from cheap execution (Haiku-tier).
The planner creates a step-by-step plan once; a pure-R dispatcher routes each step to the
correct worker without any LLM call; the evaluator decides whether the results are complete
or need a revised plan.

```{r planner, eval = FALSE}
runner <- planner_workflow(
  planner = agent(
    "planner", chat_anthropic(model = "claude-opus-4-6"),
    instructions = paste0(
      "Break the task into steps for 'researcher' and 'writer'. ",
      "Respond with one step per line as: worker_name: instruction"
    )
  ),
  workers = list(
    researcher = agent(
      "researcher", chat_anthropic(model = "claude-haiku-4-5-20251001"),
      instructions = "Research thoroughly and return structured notes."
    ),
    writer = agent(
      "writer", chat_anthropic(model = "claude-haiku-4-5-20251001"),
      instructions = "Write clearly based on the provided notes."
    )
  ),
  evaluator = agent(
    "evaluator", chat_anthropic(model = "claude-opus-4-6"),
    instructions = paste0(
      "Review the completed work against the original task. ",
      "Reply 'done' if complete, or 'replan' with a brief reason if not."
    )
  ),
  max_replans = 2L,
  max_steps   = 6L
)

result <- runner$invoke(list(
  messages = list("Write a short report on the benefits of tidy data in R.")
))

result$get("results")   # list of all worker outputs, in order
```

**How the plan is parsed**: the planner must respond with one step per line in the format
`worker_name: instruction`. Lines that do not match this format raise an error. A custom
`parse_plan` function can be supplied for alternative formats such as JSON:

```{r planner-custom-parser, eval = FALSE}
parse_json_plan <- function(text) {
  steps <- jsonlite::fromJSON(as.character(text))
  lapply(seq_len(nrow(steps)), function(i) {
    list(worker = steps$worker[[i]], instruction = steps$instruction[[i]])
  })
}

runner2 <- planner_workflow(
  planner    = agent("planner", chat_anthropic(model = "claude-opus-4-6"),
                     instructions = "Respond with a JSON array of {worker, instruction} objects."),
  workers    = list(analyst = agent("analyst", chat_anthropic())),
  parse_plan = parse_json_plan
)
```

**Cost profile** for a plan with N steps and R replanning rounds:

| Model tier | Node | Calls |
|---|---|---|
| Opus | planner | `R + 1` |
| Opus | evaluator | `R + 1` |
| Haiku | each worker | up to `N × (R + 1)` |

In a `supervisor_workflow` the manager (Opus) is called once *per step*. In a
`planner_workflow` Opus is called once *per round* regardless of plan length — a meaningful
saving for plans with many steps.

## Comparison

| Workflow | When to use |
|---|---|
| `sequential_workflow()` | Fixed pipeline: each step refines the previous output |
| `supervisor_workflow()` | Dynamic delegation: manager decides which specialist to call next |
| `debate_workflow()` | Multiple perspectives: agents challenge each other |
| `advisor_workflow()` | Quality gate: cheap model generates, expensive model approves |
| `planner_workflow()` | Cost efficiency: expensive model plans once, cheap models execute |

For anything more complex — handoffs, parallel fans, custom state channels — build the graph
directly with `state_graph()`. See `vignette("custom-graphs")`.

For guidance on model selection, context management, retry configuration, and prompt design
see `vignette("best-practices")`.
